---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, load data}
unzip("activity.zip")
data <- read.csv("activity.csv")
data$date <- as.Date(data$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r, load ggplot2}
library(ggplot2)
```
```{r, steps historgram}
steps_per_day <- aggregate(steps ~ date, data, sum, na.rm = TRUE)
ggplot(steps_per_day, aes(steps)) + 
  geom_histogram(fill="blue",binwidth=800) +
  ggtitle("Steps per day") + 
  theme_dark()
  
```

Mean of steps per day: 
```{r, steps mean}
mean(steps_per_day$steps, ra.rm=TRUE)  
```
Median of steps per day: 
```{r, steps median} 
median(steps_per_day$steps, ra.rm=TRUE)
```

## What is the average daily activity pattern?
```{r, daily_pattern}
pattern <- aggregate(steps~interval, data, mean, na.rm = TRUE)
ggplot(pattern, aes(x=interval,y=steps)) + 
  geom_line(col="blue") + 
  ggtitle("Average daily activity pattern") + 
  theme_dark()
  
```

5-min time Interval with maximum average steps taken per day:

```{r Max Avg Daily Number of Steps}
pattern$interval[pattern$steps == max(pattern$steps)]

```


## Imputing missing values
Total number of missing values:
```{r total NAs}
sum(is.na(data))
```
Imputing missing values:
```{r imput missing data}
new_data <- data
sapply(unique(data$interval),
       function(x)
         new_data[!complete.cases(new_data) & (new_data$interval == x),1] <<- pattern$steps[pattern$interval == x])
```

```{r, steps historgram with artifical data}
steps_per_day <- aggregate(steps ~ date, new_data, sum)
ggplot(steps_per_day, aes(steps)) + 
  geom_histogram(fill="blue",binwidth=800) +
  ggtitle("Steps per day with artifical data") + 
  theme_dark()
  
```

Mean of steps per day: 
```{r, steps mean with artifical data}
mean(steps_per_day$steps)  
```
Median of steps per day: 
```{r, steps median with artifical data} 
median(steps_per_day$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r, add day} 
new_data$weekday <- weekdays(new_data$date)
```
```{r, load cowplot,message=FALSE}
library(cowplot)
```

```{r, weekday amd weekend patterns}
new_data_weekday = subset(new_data, weekday %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
pattern <- aggregate(steps~interval, new_data_weekday, mean, na.rm = TRUE)
weekday <- ggplot(pattern, aes(x=interval,y=steps)) + 
    geom_line(col="blue") + 
    ggtitle("Avg activity pattern on weekday") + 
    theme_dark()

new_data_weekend = subset(new_data, weekday %in% c("Saturday", "Sunday") )
pattern <- aggregate(steps~interval, new_data_weekend, mean, na.rm = TRUE)
weekend <-ggplot(pattern, aes(x=interval,y=steps)) + 
    geom_line(col="blue") + 
    ggtitle("Avg activity pattern on weekend") + 
    theme_dark()

plot_grid(weekday,weekend)
    
```


